---
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Exploratory Data Analysis 
#### Leaving the HTML - CSS styles in the document, just in case if it is required later. 

<style>
p.blue_box {
  background-color: #4955a1;
  color: #FFFFFF;
  padding: 10px;
  border: 1px solid blue;
  margin-left: 25px;
  border-radius: 5px;
  font-style: normal;
}

p.heading {
  background-color: #a19549;
  margin: auto;
  width: 80%;
  padding: 5px;
  font-style: bold;
}

p.regular {
  margin: auto;
  padding: 5px;
}

</style>

<p class="blue_box">
Data Science projects begin with the problem at hand, and then we examine the available data. The available data should now assist us in addressing the problem. If the data itself is incorrect or insufficient, we must take a pause and either begin collecting additional data, new and relevant data, or an entirely new dataset in order to solve the problem. To determine if the data at hand is accurate, we will begin with descriptive statistics to determine the range of data values, mean, median, etc., followed by exploratory data analysis to determine the relationship between variables. Various techniques are employed during this analysis, beginning with the creation of visualizations, data cleansing to remove missing values, imputation of missing data with meaningful and close estimates, etc. </p>

<br>

<p class="heading"> <b>Problem statement: Who are the prospective customers from the available dataset who do not have health insurance and who should be contacted by our marketing team?</b> </p>

</br>

<p class="regular"> For this analysis, the data is assumed to be stored in a single data frame. But, generally, the data is spread across multiple tables in a normalized fashion (We are talking about the database normalization).  

The original dataset used for this exploratory analysis is available at the "Practical Data Science with R" author's primary [github repository](https://github.com/WinVector/PDSwR2)
</p>

```{r read_data, echo=F}
file_location <- paste0(getwd(), "/Datasets/custdata.rds")
customer_data <- readRDS(file_location)
```

- Pay special attention to the missing values in the summary

```{r summarize_data}
summary(customer_data)
```

```{r handling_missing_data}
temp <- c(1, NA, -3, 5)
is.na(temp)
```
- Now looking at the columns that have huge variation, can we explore these columns for outliers? 
- For example, the mean value of gas_usage is 41 and the 3rd quartile has 60. However, there is a max value of 570. Do you think there are any outliers in the data? 
- Plot a boxplot and see the outliers? 

```{r boxplot1}
library(ggplot2)
ggplot(customer_data) + 
  geom_boxplot(aes(x = gas_usage))

```

- What is the spread of income? Is it normally spread? Draw a histogram. 
```{r histogram}
ggplot(data=customer_data[customer_data[,"income"] <
                            min(sort(customer_data$income, decreasing = T)[1:500]), ]) + 
  geom_histogram(mapping = aes(x = income)) +
  lims(x = c(0,200000)) +
  scale_x_continuous(labels = scales::dollar_format())
```

- Density plots are another method for analyzing the distribution of univariate data  
- Use the density plot to examine the age range of the customers.  
```{r density_plot}
ggplot(customer_data) +
  geom_density(aes(x = age)) + 
  geom_text(aes(x = 2, y = 0.003), label = "xyz") 
``` 

```{r density_plot_log10}
ggplot(data=customer_data[customer_data[,"income"] <
                            min(sort(customer_data$income, decreasing = T)[1:500]), ]) + 
  geom_density(mapping = aes(x = income)) +
  scale_x_continuous(labels = scales::dollar_format()) 
```



- Create a bar chart to present the count values of various marital_statuses  
- Compare a stacked bar_plot against a side-by-side bar plot
- It would have been nice if there are proportion values to present the bar plots. This can be easily done with the dplyr and tidyr functions  

```{r bar_chart}
ggplot(customer_data) + 
  geom_bar(aes(x = marital_status, fill=health_ins), position="dodge") 
```

- We need count of each combination 

```{r aggregate_counts}
health_ins_count_by_marital_status <- aggregate(sex~marital_status+as.factor(health_ins), 
                                                data = customer_data, length)
names(health_ins_count_by_marital_status) <- c("marital_status","health_ins", "ct")
```

```{r bar_graph_counts}
ggplot(health_ins_count_by_marital_status, aes(x=marital_status, y = ct)) +
  geom_bar(aes(fill=health_ins), stat = "identity") +
  geom_text(aes(label = ct), vjust = -1)
```


- Create a bar chart to present the count values of various marital_statuses 
```{r bar_chart_desc}
ggplot(customer_data) + 
  geom_bar(aes(x = reorder(state_of_res, state_of_res, function(x) length(x)))) +
  #theme(axis.text.x = element_text(angle = 90))
  coord_flip()

```

- Compare marital status against housing type: Categorical vs. Categorical comparison 
- Facet plots with bar charts 

```{r facet_bar_plots}
ggplot(customer_data[!is.na(customer_data$housing_type),]) +
  geom_bar(aes(x = reorder(marital_status, marital_status, function(x) {length(x)}))) + 
  facet_wrap(~housing_type, scale = "free_x") +
  scale_x_discrete(name = "Marital Status") +
  coord_flip()
```

- Compare population densities across categories 

```{r density_comparison}
ggplot(data = customer_data) +
  geom_density(aes(x = age, color = marital_status, linetype = marital_status, position="fill")) + 
  scale_color_brewer(palette="Dark2")
```


- Scatter plots to identify relationships between two continuos variables. 
- Create a scatter plot between age and income to see if there is any relationship between them or they are just independent variables 

```{r scatter-plot1}
ggplot(customer_data) + 
  geom_point(aes(x = age, y = income)) + 
  ggtitle(label = "Correlation analysis between Age and Income") + 
  geom_smooth(aes(x = age, y = income), method = "auto") + 
  lims(y = c(0, 200000))
```

<p class="blue_box">
In summary, the two dimensional visualizations provide a great starting point for exploratory data analysis. The below table summarizes  </p>

Visualization | Usage 
---------| -------- 
Line plot | Shows the relationship between two continuous variables. Best when that relationship is functional, or nearly so.
Scatter plot | Shows the relationship between two continuous variables. Best when the relationship is too loose or cloud-like to be easily seen on a line plot.
Smoothing curve | Shows underly ing “average” relationship, or trend, between two continuous variables. Can also be used to show the relationship between a continuous and a binary or Boolean variable: the fraction of true values of the discrete variable as a function of the continuous variable. 
Stacked bar chart | Shows the relationship between two categorical variables (var1 and var2). Highlights the frequencies of each value of var1.
Side-by-side bar chart | Shows the relationship between two categorical variables (var1 and var2). Good for comparing the frequencies of each value of var2 across the values of var1. Works best when var2 is binary.
Filled bar chart | Shows the relationship between two categorical variables (var1 and var2). Good for comparing the relative frequencies of each value of var2 within each value of var1. Works best when var2 is binary.


- More into modeling issues later. 
