---
title: "04_Exploratory_Data_Analysis"
author: "Naga Vemprala"
date: "9/8/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

#### This R markdown demonstrates tools available to read data from various data 
#### sources. Clean and process the data for data exploration. Summarizing output.
#### Some of the concepts are repeated between "Exploratory Data Analysis" and 
#### "Data Wrangling, Data Cleaning, and Summarizing Data" 
#### Data comes in various formats and forms. 
#### Great source for datasets: https://archive.ics.uci.edu/ml/datasets.php 
#### -


``` {r}
#' ------------------------------------------------------------------------- '#
#'                       Reading and Writing data 
#' ------------------------------------------------------------------------- '#
#' 
# Reading CSV files 
fileName <- paste0(getwd(),"/../Datasets/oldCarsInsuranceData.csv")
carsData1985 <- read.csv(fileName )

```

``` {r}
# Reading Excel Tab delimited files using read.table 
fileName <- paste0(getwd(),"/../Datasets/oldCarsInsuranceData1985.xlsx")
carsData1985 <- read.table(fileName )
head(carsData1985)
```

```{r}
library(tidyverse)
# We can check for the presence of a file using file.exists function 
if (file.exists(fileName)) {
    carsData1985 <- read_csv(fileName)
    class(carsData1985)
}

``` 
``` {r} 
# Downloading files from a specified URL 
newsURL <- "https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2FDATA.gov_NGAP.xlsx"
download.file(newsURL, destfile = "test.xlsx")

```


``` {r}
# Pay attention that the github R script file we are reading is of raw data. If we 
# need to provide the URL link, download.file() downloads HTML tags as well 

url <- "https://raw.githubusercontent.com/NagaVemprala/R_Programming/main/03_ControlStructures_Functions.R"
download.file(url, destfile = "delete_ControlStructures_Functions.R")
```

``` {r}
#' Most widely used libraries for reading excel data are openxlsx and xlsx 
library(openxlsx)
carsData1985 <- read.xlsx(fileName)
head(carsData1985$make)
```

``` {r}

# An excel file with multiple sheets can be read using the sheet argument 
# Additional arguments such as startRow, skipEmptyRows helps read slice of data
fileName <- paste0(getwd(),"/../Datasets/oldCarsInsuranceData1985Split.xlsx")
carsData1985 <- carsData1985[0,]
for (sheet in c("Sheet1", "Sheet2")) {
    carsData1985 <- rbind(carsData1985, read.xlsx(fileName, 
                                                  sheet = sheet,
                                                  colNames = F,
                                                  startRow = 2))
}

head(carsData1985)
```

``` {r}
# Reading XML data (xml2 is a new library still under development.)
# There is also a XML library available. For basic XML data scraping either of 
# these two work great 
library(xml2)
newsURL <- "https://www.indiatoday.in/rss/home"
topics <- read_xml(newsURL)
xml_name(topics)
xml_children(topics)
listOfNewsLinks <- xml_children(topics) %>% xml_find_all("//link") %>% xml_text()
print(listOfNewsLinks[1:5])
```

``` {r}
library(xml2)
#librart(magrittr) #for pipe operator
url <- "http://catsalut.gencat.cat/web/.content/xml/out.xml"
doc <- read_xml(url)
#get the minicipi nodes
mun.nodes <- xml_find_all(doc, ".//municipi")
#build data.frame
df <- data.frame(
  nom.mu  = mun.nodes %>% xml_attr("nom_mu"),
  cod_mu  = mun.nodes %>% xml_attr("cod_mu"),
  nom_com = xml_find_first(mun.nodes, xpath = ".//ancestor::comarca") %>% xml_attr("nom_com"),
  cod_com = xml_find_first(mun.nodes, xpath = ".//ancestor::comarca") %>% xml_attr("cod_com"),
  nom_reg = xml_find_first(mun.nodes, xpath = ".//ancestor::regio") %>% xml_attr("nom_reg"),
  cod_reg = xml_find_first(mun.nodes, xpath = ".//ancestor::regio") %>% xml_attr("cod_reg"))
head(df)
```
``` {r}

library(xml2)
dat <- read_xml("https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2Frestaurants.xml")
dat <- xml_find_all(dat, ".//row") 
xml_attr(dat[19], "_address")

```

``` {r}

# Reading HTML data
newsURL <- "https://www.rebeccabarter.com/categories/visualization/"
library(rvest)
read_html(newsURL) 
read_html(newsURL) %>% 
    html_nodes("div") %>% 
    html_nodes("[class='title']") %>% 
    html_text()
```


``` {r}

#' ------------------------------------------------------------------------- '#
#'          Exploratory Data Analysis (EDA) - Summarizing data   
#' ------------------------------------------------------------------------- '#
#' 

#' Important steps in the EDA involves: 
#' 1. Understanding the data (Cleaning the data)
#' 2. Removing any irregularities 
#' 3. Visualizing the data to assess using a naked eye what the data offers 
#' 4. Answering "what type of preliminary tests should be performed before 
#'    running a statistical model 
#'     
fileName <- paste0(getwd(),"/Car_sales.csv")
carSales <- read.csv(fileName)

#' Handling NA values in vectors and dataframes 
#' Running summary statistics provides high-level information about the presence 
#' of NA values in the dataframe along with other meaningful descriptive statistics 
#' 
summary(carSales) # Provides summary of the complete data frame 
summary(carSales$Sales_in_thousands)

```

``` {r}
#' Role of NA in a vector of dataframe 
#' Revisit the way we create vectors first 

classScores <- c(76,65,83,91,85)
print(classScores)
# let us say that we want to remove the second score that is less than 70
# set the second element to NULL
# Create a logical vector first and then set the value to false for the position which needs to be removed 
myBoolVector <- 1:length(classScores) 
myBoolVector <- !myBoolVector %in% c(2)
myBoolVector

# Pass the logical vector as input selection 
classScores <- classScores[myBoolVector]
mode(classScores)
print(classScores)

```
``` {r}
# However, removing the elements from a list is quite straight forward 
classScoresList <- list(student1 = 76, student2=65, student3=83, student4=91, student5=85)
classScoresList
length(classScoresList)
# Now we want to remove the second element. Simply set the element to NULL
print('After removing the unnecessary element: ')
classScoresList[2] <- NULL 
classScoresList
length(classScoresList)
```

``` {r}
# What if, if we really want a list to hold NULL 
lst <- list("A", NULL, "C")
length(lst)
lst[2]
```
``` {r}
# Use a compact function to remove the NULL values 
length(lst)
lst <- compact(lst)
length(lst)
```

``` {r}
# Removing list elements using a condition 
lst <- list(NA, 0, NA, 1, 2)
lst <- lst %>% 
    discard(is.na)
lst
```
``` {r}
is_na_or_null <- function(x) {
    is.na(x) || is.null(x)
}
lst <- list(NA, 0, NULL, 1, 2)
lst %>% 
    discard(is_na_or_null)
# Opposite of discard is keep function. Elements that are required are kept rest can be discarded
```

``` {r}
# Removing rows from a dataframe containing NAs can be done using omit 
dim(carSales) 
carSales1 <- na.omit(carSales)
dim(carSales1)
```

``` {r}
# Plots helps to understand data much better than various statistical results, summary statistics 
# Three basic plots that mostly come handy are: 
# 1. Univariate 
# 2. Bivariate 
# 3. Multivariate 
# Basic tools such as annotating, providing labels, color coding enhances the aesthetics 

# The plotting related functions reside in the graphics library of the base package 
# graphics include base plots such as a simple plot, scatterplot, hist for histogram, 
# and boxplot
# Also, graphic systems are primarily 3. 
# 1. Base graphics system 
# 2. Lattice Graphics System (additional flexibility through single function approach)
# 3. ggplot2 Graphics System (Incremental approach)
#plot(carSales$Horsepower, carSales$Fuel_capacity) # Basic plot 
str(carSales)
plot(carSales$Fuel_capacity, carSales$Horsepower,
     main = "A scatter plot between Horsepower of a car against Fuel Capacity", 
     xlab = "Fuel capacity", ylab = "Horsepower")
```

``` {r}
# The next level of basic plots include Lattice system where we can add another 
# variable into our analysis (Third variable). Multiple plots for visualizing using 
# multiple predictors perspective 
carSales$EngineSizeRange <- cut(carSales$Engine_size,c(0,2,4,6,8))
carSales$EngineSizeRange <- factor(carSales$EngineSizeRange, labels = c("engine size 0-2",
                                   "engine size 2-4",
                                   "engine size 4-6",
                                   "engine size 6+"), 
                                   ordered = T)
library(lattice)
# There are other functions of xyplot such as bwplot and levelplot 
# Other beautiful plots such as violin plots and geographical plots can be drawn using them

xyplot(carSales$Fuel_capacity~carSales$Horsepower | carSales$EngineSizeRange, 
     title = "A scatter plot between Horsepower of a car against Fuel Capacity", 
     xlab = "Fuel capacity", ylab = "Horsepower")

# There are many other options lattice plots offer which can be obtained from the help 
## ?xyplot()
```

``` {r}
# Improved version of plotting can be done using ggplot 
# ggplot follows an incremental approach 
library(ggplot2)
## ?qplot()
qplot(carSales$Fuel_capacity,carSales$Horsepower,
      main = "A scatter plot between Horsepower of a car against Fuel Capacity", 
      xlab = "Fuel capacity", ylab = "Horsepower")

```

``` {r}
#' Other univariate analysis include histograms and boxplots 
#hist(carSales$Fuel_capacity)
hist(carSales$Fuel_capacity, breaks=6, 
     main="Histogram of Fuel Capacity",
     xlab = "Fuel Capacity",
     labels = T)
```

``` {r}
boxplot(#carSales$Fuel_capacity,
        carSales$Horsepower,
        main="Fuel Capacity/Horsepower", frame.plot=FALSE)

boxplot(#carSales$Fuel_capacity,
    carSales$Horsepower ~ carSales$EngineSizeRange,
    main="Fuel Capacity/Horsepower", frame.plot=FALSE)

```

``` {r}
# Scatter plot
library(car) # You can install this car package. This comes in handy while working on applied regression analysis
# You can combine boxplots and scatterplots into a single graph without much effort 
scatterplot(carSales$Fuel_capacity,carSales$Horsepower, 
            xlab = "Fuel capacity", ylab = "Horsepower", 
            main="A scatter plot between Horsepower of a car against Fuel Capacity", smooth=FALSE)
```

``` {r}
print(' ----------------------------------------- ')
```
?par()
Several parameters can only be set by a call to par():

"ask",

"fig", "fin",

"lheight",

"mai", "mar", "mex", "mfcol", "mfrow", "mfg",

"new",

"oma", "omd", "omi",

"pin", "plt", "ps", "pty",

"usr",

"xlog", "ylog",

"ylbias"