---
title: "Probability, ANOVA and Regression"
author: "Naga Vemprala"
date: "2023-03-17"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Probability distributions

- Some of the code snippets are taken from the "R Cookbook" textbook, chapter 8. 

- r<distribution name> 

- rnorm()
- dnorm() # Y-axis  
- pnorm() 
- qnorm() # x-axis 

```{r probability_functions}
var1 <- rnorm(10)
print("Random 10 normally distributed values with mean 0 and sd 1: ")
print(var1)

var1 <- rnorm(10, mean = 5, sd = 0.75) 
print("Random 10 normally distributed values with mean 5 and sd 0.75: ")
print(var1)

# The below function should return highest density.
print(dnorm(5, mean = 5, sd = 0.75))

# Relatively less densest point than the point with value of 5. 
print(dnorm(4, mean = 5, sd = 0.75))

# dnorm returns the y-axis value. 
var1 <- rnorm(1000, mean = 5, sd = 0.75) 
library(ggplot2)
ggplot(as.data.frame(var1)) + 
    geom_density(aes(x = var1))
```


```{r probability_functions_normal}
# qnorm x-axis value at a particular proportion 
qnorm(0.4, mean = 5, sd = 0.75)

# What percentage of values are less than a given input value from a normal distribution with a mean of 5 and sd of 0.75
print(pnorm(3, mean = 5, sd = 0.75)) # This proportion value should be small as the sd is small 

print(pnorm(3, mean = 5, sd = 1.5)) # This proportion value should be larger the previous one because the sd is greater than 0.75
```

```{r probability_functions_binomial}
# create a random binomial distribution of 30 elements with size of 10 and probability of success is 0.3
var2 <- rbinom(30, size = 10, prob = 0.3)
#print(var2)
hist(var2)

# What is the density of having 6 wins out of 10 trials (60% win in a given sample) with a win - rate of 30% 
dbinom(6, size = 10, prob = 0.3) # This should be a small value 

# What is the density of having 3 wins out of 10 trials (30% win in a given sample) with a win - rate of 30% 
dbinom(3, size = 10, prob = 0.3) # This should be a higher value than the previous density function. Because 3 out of 10 successes should be expected to be high from a binomial distribution with success rate of 30% out of 10 trials size. 


qbinom(0.5, size = 10, prob = 0.3)

# Binomial distributions with varying trial sizes 
sample_size <- 1:20
ggplot(data = as.data.frame(sample_size)) + 
    geom_line(aes(x = sample_size, y = dbinom(sample_size, size = 10, prob = 0.3)), color = "green") + 
    geom_line(aes(x = sample_size, y = dbinom(sample_size, size = 15, prob = 0.3)), color = "red") + 
    geom_line(aes(x = sample_size, y = dbinom(sample_size, size = 20, prob = 0.3)), color = "blue") 

```
#### The Poisson distribution is a probability distribution that is used to model the probability that a certain number of events occur during a fixed time interval when the events are known to occur independently and with a constant mean rate.

For example, suppose a given call center receives 10 calls per hour. We can use a Poisson distribution calculator to find the probability that a call center receives 0, 1, 2, 3 … calls in a given hour:

- P(X = 0 calls) = 0.00005
- P(X = 1 call) = 0.00045
- P(X = 2 calls) = 0.00227
- P(X = 3 calls) = 0.00757

**Reference:**[Poisson Distribution](https://www.statology.org/poisson-distribution-real-life-examples/)


```{r probability_functions_poisson}
# 
var3 <- rpois(12, 3)
var3

print(dpois(5, lambda = 3)) # This should be a small value 

print(dpois(4, lambda = 3)) # A higher density value should return than the previous one 

print(dpois(3, lambda = 3)) # exactly equal to 3, therefore, this is the highest density point. 

# Increase the lamdba, the poisson distribution should approach a normal distribution 

sample_size <- 1:40
ggplot(data = as.data.frame(sample_size)) + 
    geom_line(aes(x = sample_size, y = dpois(sample_size, lambda = 5)), color = "green") + 
    geom_line(aes(x = sample_size, y = dpois(sample_size, lambda = 15)), color = "red") + 
    geom_line(aes(x = sample_size, y = dpois(sample_size, lambda = 20)), color = "blue")

```

## ANOVA - Analysis of Variance 
Analysis of variance (ANOVA) is a collection of statistical models and their associated estimation procedures (such as the "variation" among and between groups) used to analyze the differences among means.

- Is there a significant difference between the average sizes of garlic bulbs treated with four distinct fertilizers?

```{r}
garlicBulbs <- read.csv(paste0(getwd(), "/Datasets/GarlicBulbSizes.csv"))
ggplot(data = garlicBulbs) + 
    geom_boxplot(aes(x = BulbSize)) +
    facet_grid(~BulbType) + 
    coord_flip()
```

```{r base_boxplot}
groupMeans <- sapply(split(garlicBulbs$BulbSize, garlicBulbs$BulbType), 
                     mean)
boxplot(BulbSize ~ BulbType, data = garlicBulbs) 
points(groupMeans, col = "red", pch = 16)
```




```{r garlic_bulbs_anova}
anovaModel <- oneway.test(garlicBulbs$BulbSize ~ garlicBulbs$BulbType)
anovaModel
```
```{r aov_tukeyHSD}
anovaModel <- aov(garlicBulbs$BulbSize ~ factor(garlicBulbs$BulbType))
plot(TukeyHSD(anovaModel))
```


 
#### Manually calculate the F- Statistic 

F = SSB / SSW 

```{r F-Statistic-Manually}
SSW <- function(x, mean_x) {
    sum( (x-mean_x)^2)
}

withInGroupVariation <- sum(mapply(FUN=SSW,
                                   split(garlicBulbs$BulbSize, garlicBulbs$BulbType),
                                   groupMeans))
withInGroupVariation <- withInGroupVariation/28 # Correcting it for degrees of freedom
betweenGroupVariation <- sum( 8 * (groupMeans - mean(garlicBulbs$BulbSize) )^2 ) 
betweenGroupVariation <- betweenGroupVariation / 3 # Correcting it for degrees of freedom 
F_statistic <- betweenGroupVariation / withInGroupVariation
p_value <- 1 - pf(F_statistic, 28, 3)
print(paste("F_statistic:", F_statistic, "& ", "p_value:", p_value))

```

#### Regression analysis 


```{r regression_analysis}
carSales <- read.csv(paste0(getwd(), "/Datasets/Car_sales.csv"))
summary(carSales)
```
 
``` {r remove_character_data}
carSales <- carSales[, -3]
```


``` {r data_processing}
# Replace NA values with the means 
imputeMeans <- function(x) {
    ifelse(is.na(x), mean(x, na.rm = T), x)
}

carSales <- as.data.frame(sapply(carSales, imputeMeans))
summary(carSales)
```

```{r correlation_analysis}
correlation_matrix <- as.data.frame(cor(carSales))
```

``` {r carSales_dataset}
carsales_model <- lm(Price_in_thousands ~ . , data = carSales)
summary(carsales_model)
```

#### Analyzing models using the AIC - Maximum Likelihood estimation

- Refer R Cookbook for forward pass and backward pass model selection.

```{r model_selection}
min.model <- lm(Price_in_thousands ~ 1, data=carSales)
fwd.model <- step(min.model, 
                  direction = "forward", 
                  scope = (Price_in_thousands ~ Sales_in_thousands + 
                    Resale_value        + 
                    Engine_size        +
                    Horsepower         +
                    Wheelbase           +
                    Width              +
                    Length             +
                    Curb_weight         +
                    Fuel_capacity      +
                    Fuel_efficiency     +
                    Power_perf_factor)
)
summary(fwd.model)
```

#### Price_in_thousands ~ Power_perf_factor + Horsepower + Engine_size + Resale_value + Curb_weight + Fuel_efficiency + Length
