---
title: "07_Plotting_Data"
author: "Naga Vemprala"
date: "9/23/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## This document provides various types of graphs that can be generated using the three available graph systems of R 

```{r simulated data}
# Generate "n" random numbers using various types of probability distributions 
# Provide the n, mean, and standard deviations, or any other required values depending on the type of distribution 

n<-1000
x1<-rnorm(n, 220,40)
x2<-rnorm(n, 220,80)
x3<-rpois(n, 35)
x4<-rnorm(n, 35,sqrt(35))
x5<-100+rexp(n, 1/25)
x6<-runif(n, 100,200)
x7<-rpois(n, 5)
x8<-round(runif(n, 10,20))

# Save the simulated data into a dataframe 
dat<-data.frame(x1,x2,x3,x4,x5,x6,x7, x8)

```

``` {r}
head(dat)
```

```{r simulated_data}
# Plot the histograms of various distributions into 2 rows and 3 columsn 
# Basic R plotting system supports this approach 
par(mfrow=c(2,3))
#par(mfcol=c(2,3))
hist(dat[, 1])
hist(dat[, 2])
hist(dat[, 3])
hist(dat[, 4])
hist(dat[, 5])
hist(dat[, 6])
```
```{r}
# Basic R plotting system supports this approach 
par(mfrow=c(2,3))
#par(mfcol=c(2,3))
# Change color; xlab 
hist(dat[, 1], main = "Distribution with mean 220, sd 40") 
# Change color; xlab
hist(dat[, 2], main = "Distribution with mean 220, sd 80")
# Set title; xlab
hist(dat[, 3], col = "green") # Set title 
# Set title; Set title 
hist(dat[, 4], xlab = "X4") # 
# Decrease the breaks to 4 
hist(dat[, 5], main = "Poisson distribution",
     col = "red",
     xlab = "X5"
     #, Add breaks here 
     )
#
hist(dat[, 6])
```
# Creating a linear regression
Housing price = 100,000 base price + 
                school-rating * 50,000 + 
                (areaInSFT * range(130,230))
                (crime rate in 1-10 * -15,000) + 
                (Businesses in the neighborhood * 5000)

``` {r}
dataSamples <- 1000
housePrice <- 100000
schoolRating <- round(runif(dataSamples, 4, 10))
areaInSft <- round(runif(dataSamples, 1000, 3800))
crimeRate <- runif(dataSamples, 1, 100)
businesses <- round(runif(dataSamples, 10, 20))
housePrice <- housePrice + 50000 * schoolRating -
    1500 * crimeRate + 
    5000 * businesses +
    #areaInSft[i in runif(1, 130,230) * areaInSft 
    sapply(areaInSft, FUN = function(x) { x * runif(1, 130,230)})
```

``` {r}
plot(areaInSft, housePrice)
```
``` {r}
# Scale the y-axis of housing price so that exponential scaling is not used as see above 
housePrice <- housePrice/100000
# Display sample housing prices to make sure the points are as expected 
housePrice[1:3]

# The position of the text is looking ugly can you please adjust it? 
plot(areaInSft, housePrice, main = "Housing Price rise due to lot size of house ",
     xlab = "Area of house",
     ylab = "Housing price")
text(2000, 9.5, "The housing price is scaled at $100,000 and this messages is\n 
           defaulted to 2000, 9 coordinates. Change it please",
          cex = 1, 
          col = "red")

```
``` {r fig.dim = c(20, 10)}
# Also, observe that the housing price scale operation "housePrice <- housePrice/100000" is removed. 
# You can control this operation using a conditional operation logic. 
# Please try it out! 

# Display sample housing prices to make sure the points are as expected 
housePrice[1:3]

# The position of the text is looking ugly can you please adjust it? 
plot(areaInSft, housePrice, main = "Housing Price rise due to lot size of house",
     xlab = "Area of house",
     ylab = "Housing price")
text(mean(areaInSft), max(housePrice)-0.25, " The housing price is scaled at $100,000 and this messages is adjusted dynamically",
          cex = 1, 
          col = "blue")


```
``` {r}
# Include a R Code to display the relationship of all the X variables on the housing price
# Hint: Use mfrow or mfcol function to create panel of plots 

```

``` {r}
# Lattice Plots 
library(lattice) 
xyplot(housePrice ~ areaInSft)
# Using xyplot, creating panel plots is easy. Just add third factor variable such as school rating 
# As shown in the next plot 
```
``` {r}
# Lattice Plots 
library(lattice) 
#xyplot(housePrice ~ areaInSft )
xyplot(housePrice ~ areaInSft, groups = schoolRating
       # Add one option at a time by uncomment the below lines and see how the 
       # additional annotations are added to the plot
       , auto.key=TRUE, col = schoolRating, columns = 7
       )
```
``` {r}
# Lattice Plots 
xyplot(housePrice ~ areaInSft | schoolRating, groups = schoolRating
       # Add one option at a time by uncomment the below lines and see how the 
       # additional annotations are added to the plot
       , auto.key=TRUE, col = schoolRating, columns = 7
       )
```
``` {r fig.dim = c(20, 10)}
# Lattice Plots 
# (Observe how the size of the plot can be modified using the fig.dim for the markdown)
library(lattice) 
schoolRating <- factor(schoolRating, labels = c("4/10 school","5/10 school","6/10 school",
                                                "7/10 school","8/10 school","9/10 school",
                                                "10/10 school"))
xyplot(housePrice ~ areaInSft | schoolRating, groups = schoolRating
       # Add one option at a time by uncomment the below lines and see how the 
       # additional annotations are added to the plot
       , auto.key=TRUE
       , col = schoolRating
       , columns = 7
       , xlab = "Area in sq. feet"
       , ylab = "House price in Thousands of dollars"
       , layout = c(4, 2)
       )
```

## qplot and ggplot are demonstrated below. 
<p>
Demonstration of qplot ggplot are provided. Using the similar examples, the preloaded diamonds dataset can be visualized. 
For practice, please replicate the steps using qplot and then continue implementing same graphs using ggplot 
</p>

``` {r}
#' ------------------------------------------------------------------- '#
#'  Grammar of Graphics can be clearly seen with qplot and ggplot 
#'  To run ggplot functions, you need to install tidyverse library. ggplot2
#'  comes with tidyverse. You can load it using: library(ggplot2)
#'  We will start with a simple scatter plot displaying the same X & Y 
#'  Columns ( areaInSft and housePrice)
#'  While working on this data, there is a placeholder to work simultaneously 
#'  on Diamonds dataset 
#' ------------------------------------------------------------------- '#
library(ggplot2)
qplot(areaInSft, housePrice) # Observe that the graph is generated using vectors.
# If data frame related columns are to be used, then data = should be used. 
# Implement the same type of graph using diamonds dataset between price & carat

```
``` {r}
#' ------------------------------------------------------------------- '#
#'  Same plot that qplot has generated is implemented using ggplot2  
#'  For practice, implement  using price & carat columns of diamonds 
#' ------------------------------------------------------------------- '#
library(ggplot2)
ggplot() + # If using diamonds, pass the data frame here. 
  geom_point(mapping = aes(x = areaInSft, y = housePrice))

```

``` {r}
# Draw a average smoothing line across points 
qplot(areaInSft, housePrice, geom = c("point", "smooth")) 
``` 

``` {r}
# ggplot expects a data frame as input. Here, both areaInSft, housePrice are two double vectors. 
# Hence, having no data frame as input is represented with an empty argument in ggplot function before ,
ggplot(,aes(areaInSft, housePrice)) + 
    geom_point() +
    geom_smooth()
```
``` {r}
print("Please draw a scatter plot between displ & hwy variables of mpg dataset with a smoothing line ")
```

``` {r}
# Facets create multiple plots between same set of variables on a certain category
# e.g., if there are 5 types of diamond cuts that are possible, the diamonds dataset can be divided into 
# five groups. Five different scatter plots can be created using facets between these groups. 
# Facets make the plot function more efficient, visually colorful, and elegant with multiple options to customize
areaInThousandsSft <- areaInSft/1000
qplot(areaInThousandsSft, housePrice, 
      facets = .~schoolRating, 
      color = schoolRating,
      geom = c("point", "smooth")
      )
```

``` {r}
# Same implementation with ggplot 
areaInThousandsSft <- areaInSft/1000
housingDF <- data.frame(schoolRating = schoolRating, 
                           areaInSft = areaInSft, 
                           crimeRate = crimeRate, 
                           businesses = businesses, 
                           areaInThousandsSft = areaInThousandsSft, 
                           housePrice = housePrice)
ggplot(housingDF, aes(areaInThousandsSft, housePrice)) + 
    geom_point(aes(color = schoolRating)) +
    facet_grid(.~schoolRating) + 
    geom_smooth()
```

``` {r}
# Same implementation with ggplot in more elegant fashion by giving each of the type of 
# school raing a separate color. Observe the "aesthetics" enforced onto geom_point 
# geom_point uses the aes function to provide aesthetics to the points. 
areaInThousandsSft <- areaInSft/1000
housingDF <- data.frame(schoolRating = schoolRating, 
                           areaInSft = areaInSft, 
                           crimeRate = crimeRate, 
                           businesses = businesses, 
                           areaInThousandsSft = areaInThousandsSft, 
                           housePrice = housePrice)
ggplot(housingDF, aes(areaInThousandsSft, housePrice)) + 
    geom_point(color = schoolRating) +
    facet_grid(.~schoolRating) + 
    geom_smooth()
```

``` {r}
# Complete graph implementation with axis labels and title  
areaInThousandsSft <- areaInSft/1000
housingDF <- data.frame(schoolRating = schoolRating, 
                           areaInSft = areaInSft, 
                           crimeRate = crimeRate, 
                           businesses = businesses, 
                           areaInThousandsSft = areaInThousandsSft, 
                           housePrice = housePrice)
ggplot(housingDF, aes(areaInThousandsSft, housePrice)) + 
    geom_point(color = schoolRating) +
    facet_grid(.~schoolRating) + 
    geom_smooth() +
    #labs(x = "Area of house in thousands of Sq. ft ",
    #     y = "House price in $100,000 ",
    #     title = " Exploratory Data Analysis of Housing prices ") +
    #theme(plot.title = element_text(hjust = 0.5)) +
    
    #'  -----------------------------------------------  '#
    #'   Either of the options work to update the axes   '#
    #'  -----------------------------------------------  '#
    xlab("Area of house in thousands of Sq. ft ") +
    ylab("House price in $100,000 ") +
    ggtitle(" Exploratory Data Analysis of Housing prices " ) +
    theme(plot.title = element_text(hjust = 0.5))
```

# Please practice the last set of graphs using the diamonds and mpg data frames.