---
title: "GGPLOT2"
author: "Naga Vemprala"
date: '2023-02-05'
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Grammar of Graphics (Layered Grammar of Graphics) using ggplot2 library 

- How to graphically describe a story concisely. Good grammar aids in effectively communicating the moral of a story. The same holds true for an effective visualization.   

The general syntax of a ggplot graph looks like below: 

``` {r chunk1, eval = F}
ggplot(data = [DATA], mapping = aes([MAPPING]) +  
         [GEOM_FUNCTION]( mapping = aes([MAPPING]), 
                          stat = [STAT], 
                          position = [POSITION]) +  
         [FACET_FUNCTION]() +  
         [COORDINATE_FUNCTION]() +  
         labs(title = [TITLE],  
              x = [X AXIS LABEL],  
              y = [Y AXIS LABEL]) + 
         [SCALE_FUNCTION] + 
         [THEME_FUNCTION]
              
```

Required reading: [Data Visualization](https://r4ds.had.co.nz/data-visualisation.html)

Let's first explore the mpg dataset. 

```{r explore}
library(ggplot2)
summary(mpg)
```

```{r basic_plot}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy))
```

?geom_point() to see what other options we have here. 

geom_point() understands the following aesthetics (required aesthetics are in bold):

- **x**
- **y**
- alpha 
- colour 
- fill
- group
- shape
- size
- stroke

```{r geom_point_additions}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, 
                           color = class, 
                           shape = drv))
```

**Observe that the aesthetics that I have used here are passed in the aes function itself but not outside of the aes. If we want to control the aesthetics using the variables of the data object itself, then we must pass the aesthetics related arguments within the aes itself. We will look into another example of bar plot below: 

```{r bar_plot_with_wrong_color}
ggplot(data = mpg) + 
  geom_bar(mapping = aes(x = manufacturer, color = "blue")) # This is not the expected behavior 
```

```{r bar_plot_with_right_color}
ggplot(data = mpg) + 
  geom_bar(mapping = aes(x = manufacturer), color = "blue") # The color is for the outline and fill is to fill inside 
```

```{r bar_plot_with_right_fill}
ggplot(data = mpg) + 
  geom_bar(mapping = aes(x = manufacturer), fill = "blue") # The color is for the outline and fill is to fill inside 
```

```{r boxplot1}
ggplot(data = mpg) + 
  geom_boxplot(mapping=aes(x=hwy), position = "dodge2", outlier.colour = "red") # The default statistics of geom_boxplot is statistics. It is embedded in the stat parameter

```

```{r stacking_geoms}
ggplot(data = mpg) +
  geom_bar(mapping = aes(x = class)) + 
  geom_text(aes(x = class, label = ..count..), stat = "count", color = "white", vjust = 1) 
```
### Scale controls the legends and axis 
### Scale has three properties controlled by three arguments, name, breaks, labels
- Name is the title of the respective axis or legend 
- There are various scale functions and each function is makde up of multiple words separated by "_" 
- 1st word is scale, 2nd word is primary attribute color, shape, or x/y axis, and the 3rd word is the name of the scale, such as continuous, discrete, brewer 

e.g., 
scale_colour_brewer()
scale_x_sqrt()


```{r scales1}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, 
                           color = class, 
                           shape = drv)) + 
  scale_x_continuous(name = "Using \"hwy\" variable for Gas Mileage") 
```

```{r scales2}
ggplot(data = mpg, mapping = aes(x = class)) +
  geom_bar(aes(fill=drv)) + 
  geom_text(aes(label = ..count..), stat = "count", color = "white", vjust = 1) +
  scale_fill_hue(h = c(180, 300))
```

Lastly, let's set the legend name. Legend here is used on a discreet values and for this bar_plot, we will set a legend title as name = "Legend for Drivetrain"


```{r scales3}
ggplot(data = mpg, mapping = aes(x = class)) +
  geom_bar(aes(fill=drv)) + 
  geom_text(aes(label = ..count..), stat = "count", color = "white", vjust = 1) +
  scale_fill_hue(h = c(180, 300), name = "Legend for Drivetrain")
```

```{r scales4}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, 
                           color = class, 
                           shape = drv)) + 
  scale_x_continuous(name = "Using \"hwy\" variable for Gas Mileage") +
  scale_color_discrete(name = "My legend",
    limits = c("2seater", "compact", "midsize", "minivan", 
                                  "pickup", "subcompact", "suv"), 
    breaks = c("2seater", "compact", "midsize", "minivan", "pickup", "subcompact"
               , "suv"),
    labels = c("Two Seater", "Compact", "Mid-Size", "Minivan", "Pickup", 
               "Sub-Compact", "SUV"))
```

```{r scales_limit}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, 
                           color = class, 
                           shape = drv)) + 
  lims(x = c(1,6), 
       y = c(10, 40))
```

```{r facets1}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, 
                           color = class, 
                           shape = drv)) +
  facet_wrap(vars(manufacturer))
```

### Facets 
facet_wrap(): “wraps” a 1d ribbon of panels into 2d.
facet_grid() lays out plots in a 2d grid

```{r facets2-variation1}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, 
                           color = class)) +
  facet_grid(.~drv)
```

```{r facets2-variation2}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, 
                           color = class)) +
  facet_grid(drv ~.)
```


```{r facets3-two-variables}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, 
                           color = class)) +
  facet_grid(drv ~ cyl)
```


```{r abline}
model <- lm(hwy~displ, mpg)
ggplot(mpg, aes(displ, hwy)) + 
  geom_abline(slope = model$coefficients[[2]], intercept = model$coefficients[[1]]) +
  facet_wrap(~cyl)
```
```{r coordinate1-base}
p <- ggplot(data = mpg) +
  geom_bar(mapping = aes(x = class, fill = drv), 
           color="black",
           linetype = 2,
           position="dodge")
print(p)
```

### Flip the coordinate 
```{r}
p + coord_flip()
```

### Display the same graph on a polar coordinate system 

```{r coordinate2}
ggplot(data = mpg) +
  geom_bar(mapping = aes(x = class, fill = drv), 
           color="black",
           linetype = 2) +
  coord_polar()
```

